Extended Data

### 2021-11-15 doc/manuscript/Dow_manuscript_Nature.Rmd
Harvard forest is not mid-Atlantic

**As the climate changes, warmer spring temperatures are causing earlier leaf-out[@jeong_phenology_2011;@keenan_net_2014; @buermann_widespread_2018;@finzi_carbon_2020;@stuble_plant_2021;@delgado_differences_2020] and commencement of net carbon dioxide (CO~2~) sequestration[@keenan_net_2014;@finzi_carbon_2020] in temperate deciduous forests, resulting in a tendency towards increased growing season length[@menzel_growing_1999;@menzel_european_2006;@ibanez_forecasting_2010;@jeong_phenology_2011;@finzi_carbon_2020;@stuble_plant_2021] and annual CO~2~ uptake[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014; @crabbe_extreme_2016; @zhou_explaining_2016; @fu_maximum_2019;@finzi_carbon_2020].**
**However, less is known about how spring temperatures affect  tree stem growth, which sequesters carbon in wood that has a long residence time in the ecosystem [@xue_global_2017;@russell_residence_2014].**
**Using dendrometer band measurements from `r n_dendro_trees` trees across two deciduous mid-Atlantic forests, we show that warmer spring temperatures consistently shifted the woody growth of deciduous trees earlier but had no consistent effect on peak growing season length, maximum daily growth rates, or annual growth.**
**The latter finding was confirmed on the centennial scale by `r n_chronologies` tree-ring chronologies from `r n_cores_sites` forests across eastern North America, where annual growth was far more sensitive to temperatures during the peak growing season than in the spring.**
**These findings imply that any extra CO~2~ sequestered in years with warmer springs[@keeling_increased_1996;@dragoni_evidence_2011;@crabbe_extreme_2016] is not allocated to long-lived woody biomass, where it could have a substantial and lasting impact on the forest C balance. **
**Such empirical results contradict current projections from global carbon cycle models[@buermann_widespread_2018; @zani_increased_2020;@keenan_net_2014;@ahlstrom_robustness_2012], and imply that warming spring temperatures are unlikely to increase the woody productivity or strengthen the carbon sink of temperate deciduous forests.**

>>>>

**As the climate changes, warmer spring temperatures are causing earlier leaf-out[@jeong_phenology_2011;@keenan_net_2014; @buermann_widespread_2018;@finzi_carbon_2020;@stuble_plant_2021;@delgado_differences_2020] and commencement of net carbon dioxide (CO~2~) sequestration[@keenan_net_2014;@finzi_carbon_2020] in temperate deciduous forests, resulting in a tendency towards increased growing season length[@menzel_growing_1999;@menzel_european_2006;@ibanez_forecasting_2010;@jeong_phenology_2011;@finzi_carbon_2020;@stuble_plant_2021] and annual CO~2~ uptake[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014; @crabbe_extreme_2016; @zhou_explaining_2016; @fu_maximum_2019;@finzi_carbon_2020].**
**However, less is known about how spring temperatures affect tree stem growth, which sequesters carbon in wood that has a long residence time in the ecosystem [@xue_global_2017;@russell_residence_2014].**
**Using dendrometer band measurements from `r n_dendro_trees` trees across two forests, we show that warmer spring temperatures consistently shifted the woody growth of deciduous trees earlier but had no consistent effect on peak growing season length, maximum daily growth rates, or annual growth.**
**The latter finding was confirmed on the centennial scale by `r n_chronologies` tree-ring chronologies from `r n_cores_sites` forests across eastern North America, where annual growth was far more sensitive to temperatures during the peak growing season than in the spring.**
**These findings imply that any extra CO~2~ sequestered in years with warmer springs[@keeling_increased_1996;@dragoni_evidence_2011;@crabbe_extreme_2016] is not allocated to long-lived woody biomass, where it could have a substantial and lasting impact on the forest C balance. **
**Such empirical results contradict current projections from global carbon cycle models[@buermann_widespread_2018; @zani_increased_2020;@keenan_net_2014;@ahlstrom_robustness_2012], and imply that warming spring temperatures are unlikely to increase the woody productivity or strengthen the carbon sink of temperate deciduous forests.**
